Equivalence Testing and the Second Generation P-Value

### 2018-08-17 sgpv_vs_TOST.Rmd
fixed typo

To test predictions most researchers predominantly rely on null-hypothesis tests. 
This statistical approach can be used to examine whether observed data is sufficiently surprising under the null hypothesis to reject an effect size of zero. 
Null-hypothesis tests have an important limitation, in that this procedure can only reject the hypothesis that there is no effect, while scientists should also be able to provide statistical support for *equivalence*. 
When testing for equivalence researchers aim to examine whether the effect is practically zero, or that there is no meaninful effect. 
By specifying a range around the null hypothesis of values that are deemed practically equivalent to the absence of an effect (i.e., 0 ± 0.3) the observed data can be compared against an *equivalence range* and researchers can test if a meaningful effect is absent (e.g., Hauk & Anderson, 1985; Kruschke, 2018; Metzler, 1974; Rogers, Howard, & Vessey, 1993; Serlin & Lapsey, 1985; Spiegelhalter, Freedman, & Parmar, 1994; Wellek, 2010; Westlake, 1972).

@@

One weakness of the SGPV is its reliance on the 'small sample correction', where the SGPV is set to 0.5 whenever the ratio of the confidence interval width to the equivalence range exceeds 2:1 and the CI overlaps with the upper and lower bounds. This exception to the normal calculation of the SGPV is introduced to prevent misleading values. Without this correction it is possible that a confidence interval is extremely wide, and an equivalence range is extremely narrow, which without the correction would lead to a very low value for the SGPV, which as Blume et al (2018, p. 7) seem to suggest that 'the data favor alternative hypotheses', even when the data is just extremely inaccurate compared to the width of the equivalence range. Although setting the SGPV to 0.5 whenever the ratio of the confidence interval width to the equivalence range exceeds 2:1 is necessary to prevent misleading results, it leads to a range of situations where the SGPV is set to 0.5, while the *p*-value from the TOST procedure continues to differentiate (see Figure 6). An important benefit of equivalence tests is that is does not need such a correction to prevent misleasing results.

>>>>

To test predictions most researchers predominantly rely on null-hypothesis tests. 
This statistical approach can be used to examine whether observed data is sufficiently surprising under the null hypothesis to reject an effect size of zero. 
Null-hypothesis tests have an important limitation, in that this procedure can only reject the hypothesis that there is no effect, while scientists should also be able to provide statistical support for *equivalence*. 
When testing for equivalence researchers aim to examine whether the effect is practically zero, or that there is no meaningful effect. 
By specifying a range around the null hypothesis of values that are deemed practically equivalent to the absence of an effect (i.e., 0 ± 0.3) the observed data can be compared against an *equivalence range* and researchers can test if a meaningful effect is absent (e.g., Hauk & Anderson, 1985; Kruschke, 2018; Metzler, 1974; Rogers, Howard, & Vessey, 1993; Serlin & Lapsey, 1985; Spiegelhalter, Freedman, & Parmar, 1994; Wellek, 2010; Westlake, 1972).

@@

One weakness of the SGPV is its reliance on the 'small sample correction', where the SGPV is set to 0.5 whenever the ratio of the confidence interval width to the equivalence range exceeds 2:1 and the CI overlaps with the upper and lower bounds. This exception to the normal calculation of the SGPV is introduced to prevent misleading values. Without this correction it is possible that a confidence interval is extremely wide, and an equivalence range is extremely narrow, which without the correction would lead to a very low value for the SGPV, which as Blume et al (2018, p. 7) seem to suggest that 'the data favor alternative hypotheses', even when the data is just extremely inaccurate compared to the width of the equivalence range. Although setting the SGPV to 0.5 whenever the ratio of the confidence interval width to the equivalence range exceeds 2:1 is necessary to prevent misleading results, it leads to a range of situations where the SGPV is set to 0.5, while the *p*-value from the TOST procedure continues to differentiate (see Figure 6). An important benefit of equivalence tests is that is does not need such a correction to prevent misleading results.
